### 2019-04-02 papers/journal/encoding.tex
SMT encoding with Constraints Tag

\subsubsection{Constraints on presence, activity of the molecule, and pairing matrix}
%
We need the following constraints \texttt{EdgeC},
\texttt{ActivityC} and \texttt{PairingC} to encode the basic structure of VTSs.
%
For an edge to exist it should have one molecule present. 
%
\begin{align*}
  \texttt{PresentEdge} = \bigwedge\limits_{i,j,q} (\bigvee_m e_{i,j,q,m} )\limplies e_{i,j,q}
\end{align*}
The edge labels are subset of the node labels of the source and target nodes.
\begin{align*}
  \texttt{LabelEdge} = \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} \limplies (n_{i,m} \land n_{j,m} )
\end{align*}
Self edges are not allowed. 
\begin{align*}
  \texttt{SelfEdge} = \bigwedge\limits_{i,q} \neg e_{i,i,q}
\end{align*}

We use \texttt{EdgeC} to collect all the edge related constraints. 
%
\texttt{EdgeC} states that each edge has at least one molecule,
 edge labels are consistent with node labels and there are no self loops.    
\begin{align*}
  \texttt{EdgeC} = \texttt{PresentEdge} \land \texttt{SelfEdge} \land \texttt{LabelEdge} 
\end{align*}

If a molecule is active on an edge, it should be present on the edge.
%
\begin{align*}
  \texttt{ActivityMol} = \bigwedge\limits_{i,j,q,m} b_{i,j,q,m} \limplies e_{i,j,q,m}\label{eq:f1}
\end{align*}
A molecule should be present to be active on a node.  
\begin{align*}
    \texttt{PresentMol} = \bigwedge\limits_{i,m} a_{i,m} \limplies n_{i,m}
\end{align*}

We use \texttt{ActivityC} to collect all the molecule activity related constraints. 
\texttt{ActivityC} states that active molecule are present.
\begin{align*}
  \texttt{ActivityC} = \texttt{ActivityMol} \land \texttt{PresentMol} 
\end{align*}

We divide the pairing matrix into Q-SNAREs and R-SNAREs blocks and set the dependency bits of Q-SNAREs to Q-SNAREs (and similarly R-SNAREs to R-SNAREs) in the pairing matrix to be 0's.
%It reduces the search space.
% Condition on $p_{kk'}$. \ashu{why???} .
%\begin{align}
%  \bigwedge\limits_{(x < M/2 \, \land  \, y < M/2) \lor  (x >= M/2 \land y >= M/2)} \neg p(x,y)
%  \tag{V6}\label{eq:c3}
%\end{align}

@@

\subsubsection{Well-fused constraints}
Constraint \texttt{Fusion1} encodes that each edge must fuse with
its destination node.
%
Constraint \texttt{Fusion2} encodes that each edge should not
be able to potentially fuse with any node other than its destination node.
\begin{align*}
\texttt{Fusion1} =\;&
\bigwedge\limits_{i,j,q} e_{i,j,q} \limplies
\bigvee_{{m_1,m_2,m_3,m_4}} (\Land_{l=1}^4 ( b_{i,j,q,m_l} \lor a_{j,m_l} ) \land 
\Lor\limits_{l=1}^4 b_{i,j,q,m_l} \land\\
& \qquad \qquad \qquad \qquad \Lor\limits_{l=1}^4 a_{j,m_l}\land p_{\{m_1,m_2,m_3,m_4\}})
\\
\texttt{Fusion2} =\;&
\bigwedge\limits_{\mathclap{i,j,q,m_1,m_2,m_3, l \in \{1,..,3\}}} b_{i,j,q,m_1} \land ..\land b_{i,j,q,m_l} \limplies \\
& \hspace{2cm} \neg 
\bigvee_{\mathclap{j \neq j^{\prime}, m_{l+1}^{\prime},..,m_{4}^{\prime}}} ( a_{j^{\prime},m_{l+1}^{\prime}} \land .. \land a_{j^{\prime},m_4^{\prime}} \land p_{\{m_1,..,m_l,m^{\prime}_{l+1},..,m^{\prime}_4\}})
\\
\end{align*}

% Fusion rules consist of two different mechanisms.
% \begin{enumerate}
% \item  A \textbf{pairing mechanism} which determines compatible Q-R pairs on vesicles and compartments that can cause fusion.
% \item \textbf{regulatory mechanisms} on the edges and nodes (possibly distinct) which regulates molecules activity based on the presence/absence of other molecules on the corresponding node or edge.
% \end{enumerate}
% Boolean constraint~\eqref{eq:fuse1} and~\eqref{eq:fuse2} are used to encode well-fused
% property.

@@

\subsubsection{Regulation on nodes and edges (Activity function)}
We are considering several models that differ in constraints on
the activity of molecules.
%
%We will present~\eqref{eq:ann}-~\eqref{eq:aep} that encodes
%the varying constraints.
%
The activity of the molecules are controlled by activity functions.
%
We will have encoded five variation of the boolean functions based on the biological discussion.
%
The concrete functions encoding is presented below.
%
We will use \texttt{NodeFun}$_m$ and \texttt{EdgeFun}$_m$ for node and edge regulations respectively.

The following constraint encodes no conditions on activities on nodes,
i.e., all the present molecules on the nodes are active.
\begin{align*}
\texttt{NodeFunOn} =& \bigwedge\limits_{i,m} n_{i,m} = a_{i,m}  
\end{align*}
The following constraint encodes that activity of a molecule $m$ on the node is
defined by a Boolean function $\nodef_m$ of presence of molecules present on that node.
\begin{align*}
\texttt{NodeFunBf} =& \bigwedge\limits_{i,m} n_{i,m} \limplies a_{i,m} =  \nodef_m (n_{i,1},\dots,n_{i,\mu}) 
\end{align*}
The following constraint encodes no conditions on activities on edges,
i.e., all the present molecules on the edges are active.
\begin{align*}
 \texttt{EdgeFunOn} =& \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} = b_{i,j,q,m}
\end{align*}
The following constraint encodes that activity of a molecule $m$ on the edge is
defined by a Boolean function $\edgef_m$ of presence of molecules present on that edge.
\begin{align*}
 \texttt{EdgeFunBf} =& \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} \limplies b_{i,j,q,m} = \edgef_k(e_{i,j,q,1}, .., e_{i,j,q,\mu} )
\end{align*}
%
The following constraint encodes that the activity of the molecules on
edges is defined by inhibition by other molecules based on the pairing
matrix. 
\begin{align*}
\texttt{EdgeFunPairing} =&	
\bigwedge\limits_{i,j,q, m}  [e_{i,j,q,m} \land   \\
      \bigwedge_{m_l \neq m,l \in \{1,..,3\}} e_{i,j,q,m'} \land p_{m, m_l}] \limplies 
	(\neg b_{i,j,q,m} \land \bigwedge_{\mathclap{m_l \neq m,l, p_{m, m_l} \in \{1,..,3\}}} \neg b_{i,j,q,m_l})
\end{align*}
%
\subsubsection{Constraints for stability condition}
%
We use $m$-reachability to encode the stability condition in VTSs.
%
The following constraint recursively encodes that node $j$ is $m$-reachable from node $i$ in less than $p$ steps
if either there is a direct edge between $i$ and $j$ with $m$ present on the edge or there is a edge between $i^{\prime \prime}$ and
$(i \neq i^{\prime \prime})$ with $m$ present, and j is $m$-reachable from $i'$ in less than $p-1$ steps.
%
\begin{align*}
  \texttt{Paths}(r) &= \bigwedge\limits_{i,j,m,p} r_{i,j,m,p} \limplies (\bigvee_{q} \, e_{i,j,q,m} \lor \bigvee_{i\neq i^{\prime}} ( \, \bigvee_{q} e_{i,i^{\prime},q,m}) \land r_{i^{\prime},j,m,p-1} )
\end{align*}
Now we can encode stability using the reachability variables.
and say if there is an $m$-edge between $i$ and $j$, there is
$m$-reachable path from $j$ to $i$.
\begin{align*}
 \texttt{Loop}(r) &= \bigwedge\limits_{i,j,m} (\bigvee_{q} e_{i,j,q,m}) \limplies r_{j,i,m,\nu}
\end{align*}

% Boolean variable
% F3 and F2 are used to model these constraints.

@@

\subsubsection{$k$-connectivity constraints}
To check whether $k$-connected is a necessary condition, we remove (drop) $k-1$ edges from the graph and if it
disconnects the graph, and we get an assignment. We have a graph that is not  $k$-connected.
%
Constraints \texttt{Fail}(d,k), \texttt{FReach}(d,r') and \texttt{Connected}(r')
encode the relevant constraints for reachability
in the modified VTS. 
%
The following constraints encode that only
existing edges can be dropped and exactly $k-1$ edges are dropped.
\begin{align*}
\texttt{Fail}(d,k) = & 
\bigwedge\limits_{i,j,q} d_{i,j,q} \limplies e_{i,j,q}  \land 
\sum_{i,j,q} d_{i,j,q} = k-1\\
\end{align*}
We need to define reachability in the modified VTS, therefore we use
a new variable $r'_{i,j}$ to encode reachability from $i$ to $j$.
In the following constraint, we encode $r'_{i,j}$ is true if there is an
edge $(i,\_,j)$ and it is not dropped, or there is a node
$i^{\prime}$ such that, there is an edge $(i,\_,i^{\prime})$ which is
not dropped and $r'_{i',j}$ is true.
\begin{align*}
\texttt{FReach}(d,r') = &  \bigwedge\limits_{i,j}  [\bigvee_{q} (e_{i,j,q} \land  \neg d_{i,j,q}) \lor  (\bigvee_{i' \neq i}  r^{\prime}_{i',j} \land  \bigvee_{q} (e_{i,i',q} \land \neg d_{i,i',q}) ] \limplies r^{\prime}_{i,j}  
\end{align*}
Since we search for disconnected modified VTS,
the following constraint encodes that there are 
unreachable pair of nodes in the underlying undirected graph.
% nodes such that
% there is no path between them .
\begin{align*}
   \texttt{Connected}(r')  = & \bigvee\limits_{i,j} \neg (r^{\prime}_{i,j} \lor r^{\prime}_{j,i})
\end{align*}

% We can go down or up using -C \_ option.  \newline

>>>>

\subsubsection{Constraints on presence, activity of the molecule, and pairing matrix}
%
We need the following constraints \texttt{EdgeC},
\texttt{ActivityC} and \texttt{PairingC} to encode the basic structure of VTSs.
%
For an edge to exist it should have one molecule present. 
%
\begin{align*}
  \texttt{PresentE} = \bigwedge\limits_{i,j,q} (\bigvee_m e_{i,j,q,m} )\limplies e_{i,j,q}
\end{align*}
The edge labels are subset of the node labels of the source and target nodes.
\begin{align*}
  \texttt{LabelE} = \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} \limplies (n_{i,m} \land n_{j,m} )
\end{align*}
Self edges are not allowed. 
\begin{align*}
  \texttt{SelfE} = \bigwedge\limits_{i,q} \neg e_{i,i,q}
\end{align*}
\noindent We define \texttt{EdgeC} to collect all the edge related constraints. 
%
\texttt{EdgeC} states that each edge has at least one molecule,
 edge labels are consistent with node labels and there are no self loops.    
\begin{align*}
  \texttt{EdgeC} = \texttt{PresentE} \land \texttt{LabelE}  \land \texttt{SelfE} 
\end{align*}
\noindent If a molecule is active on an edge, it should be present on the edge.
%
\begin{align*}
  \texttt{ActiveMolE} = \bigwedge\limits_{i,j,q,m} b_{i,j,q,m} \limplies e_{i,j,q,m}\label{eq:f1}
\end{align*}
A molecule should be present to be active on a node.  
\begin{align*}
    \texttt{ActiveMolN} = \bigwedge\limits_{i,m} a_{i,m} \limplies n_{i,m}
\end{align*}
\noindent We define \texttt{ActivityC} to collect all the molecule activity related constraints. 
\texttt{ActivityC} states that all active molecules are present.
\begin{align*}
  \texttt{ActivityC} = \texttt{ActiveMolE} \land \texttt{ActiveMolN} 
\end{align*}
\noindent We divide the pairing matrix into Q-SNAREs and R-SNAREs blocks and set the dependency bits of Q-SNAREs to Q-SNAREs and R-SNAREs to R-SNAREs in the pairing matrix to be all 0's.
%It reduces the search space.
% Condition on $p_{kk'}$. \ashu{why???} .
%\begin{align}
%  \bigwedge\limits_{(x < M/2 \, \land  \, y < M/2) \lor  (x >= M/2 \land y >= M/2)} \neg p(x,y)
%  \tag{V6}\label{eq:c3}
%\end{align}

@@

\subsubsection{Well-fused constraints}
Constraint \texttt{Fusion1} encodes that each edge must fuse with
its destination node.
%
Constraint \texttt{Fusion2} encodes that each edge should not
be able to potentially fuse with any node other than its destination node.
\begin{align*}
\texttt{Fusion1} =\;&
\bigwedge\limits_{i,j,q} e_{i,j,q} \limplies
\bigvee_{{m_1,m_2,m_3,m_4}} (\Land_{l=1}^4 ( b_{i,j,q,m_l} \lor a_{j,m_l} ) \land 
\Lor\limits_{l=1}^4 b_{i,j,q,m_l} \land\\
& \qquad \qquad \qquad \qquad \Lor\limits_{l=1}^4 a_{j,m_l}\land p_{\{m_1,m_2,m_3,m_4\}})
\\
\texttt{Fusion2} =\;&
\bigwedge\limits_{\mathclap{i,j,q,m_1,m_2,m_3, l \in \{1,..,3\}}} b_{i,j,q,m_1} \land ..\land b_{i,j,q,m_l} \limplies \\
& \hspace{2cm} \neg 
\bigvee_{\mathclap{j \neq j^{\prime}, m_{l+1}^{\prime},..,m_{4}^{\prime}}} ( a_{j^{\prime},m_{l+1}^{\prime}} \land .. \land a_{j^{\prime},m_4^{\prime}} \land p_{\{m_1,..,m_l,m^{\prime}_{l+1},..,m^{\prime}_4\}})
\\
\end{align*}
We define \texttt{FusionC} to collect all the fusion related constraints.
% 
\texttt{FusionC} states that every edge in the graph is well-fused.
\begin{align*}
\texttt{FusionC} = \texttt{Fusion1} \land \texttt{Fusion2} 
\end{align*}
% Fusion rules consist of two different mechanisms.
% \begin{enumerate}
% \item  A \textbf{pairing mechanism} which determines compatible Q-R pairs on vesicles and compartments that can cause fusion.
% \item \textbf{regulatory mechanisms} on the edges and nodes (possibly distinct) which regulates molecules activity based on the presence/absence of other molecules on the corresponding node or edge.
% \end{enumerate}
% Boolean constraint~\eqref{eq:fuse1} and~\eqref{eq:fuse2} are used to encode well-fused
% property.

@@

\subsubsection{Regulation on nodes and edges (Activity function)}
We are considering several models of VTS that differ in types of regulations on the molecule both on the node and on the edge.  
%constraints on the activity of molecules.
%
%We will present~\eqref{eq:ann}-~\eqref{eq:aep} that encodes
%the varying constraints.
%
The activity of the molecules are controlled by activity functions.
%
We will have encoded five variation of the Boolean functions based on the biological discussion.
%
The concrete functions encoding is presented below.
%
We will use \texttt{NodeFun}$_m$ and \texttt{EdgeFun}$_m$ for node and edge regulations respectively.
%

The following constraint encodes no conditions on activities on nodes,
i.e., all the present molecules on the nodes are active.
\begin{align*}
\texttt{NodeFunOn} =& \bigwedge\limits_{i,m} n_{i,m} = a_{i,m}  
\end{align*}
The following constraint encodes that activity of a molecule $m$ on the node is
defined by a Boolean function $\nodef_m$ of presence of molecules present on that node.
\begin{align*}
\texttt{NodeFunBf} =& \bigwedge\limits_{i,m} n_{i,m} \limplies a_{i,m} =  \nodef_m (n_{i,1},\dots,n_{i,\mu}) 
\end{align*}
The following constraint encodes no conditions on activities on edges,
i.e., all the present molecules on the edges are active.
\begin{align*}
 \texttt{EdgeFunOn} =& \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} = b_{i,j,q,m}
\end{align*}
The following constraint encodes that activity of a molecule $m$ on the edge is
defined by a Boolean function $\edgef_m$ of presence of molecules present on that edge.
\begin{align*}
 \texttt{EdgeFunBf} =& \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} \limplies b_{i,j,q,m} = \edgef_k(e_{i,j,q,1}, .., e_{i,j,q,\mu} )
\end{align*}
%
The following constraint encodes that the activity of the molecules on
edges is defined by inhibition by other molecules based on the pairing
matrix. 
\begin{align*}
\texttt{EdgeFunPairing} =&	\begin{aligned}[t]
\bigwedge\limits_{i,j,q, m}  [e_{i,j,q,m} \land \bigwedge_{m_l \neq m,l \in \{1,..,3\}} e_{i,j,q,m'} \land p_{m, m_l}] \limplies\\
	(\neg b_{i,j,q,m} \land \bigwedge_{\mathclap{m_l \neq m,l, p_{m, m_l} \in \{1,..,3\}}} \neg b_{i,j,q,m_l})\end{aligned}\\
\end{align*}
%
We denote the particular choice of node and edge function combined by \texttt{ActivityFun}.
\begin{align*}
\texttt{ActivityFun} = \texttt{NodeFun}_m \land \texttt{EdgeFun}_m
\end{align*}
\subsubsection{Constraints for stability condition}
%
We use $m$-reachability to encode the stability condition in VTSs.
%
The following constraint recursively encodes that node $j$ is $m$-reachable from node $i$ in less than $p$ steps
if either there is a direct edge between $i$ and $j$ with $m$ present on the edge or there is a edge between $i^{\prime \prime}$ and
$(i \neq i^{\prime \prime})$ with $m$ present, and j is $m$-reachable from $i'$ in less than $p-1$ steps.
%
\begin{align*}
  \texttt{Paths}(r) &= \bigwedge\limits_{i,j,m,p} r_{i,j,m,p} \limplies (\bigvee_{q} \, e_{i,j,q,m} \lor \bigvee_{i\neq i^{\prime}} ( \, \bigvee_{q} e_{i,i^{\prime},q,m}) \land r_{i^{\prime},j,m,p-1} )
\end{align*}
Now we can encode stability using the reachability variables.
and say if there is an $m$-edge between $i$ and $j$, there is
$m$-reachable path from $j$ to $i$.
\begin{align*}
 \texttt{Loop}(r) &= \bigwedge\limits_{i,j,m} (\bigvee_{q} e_{i,j,q,m}) \limplies r_{j,i,m,\nu}
\end{align*}
We define \texttt{StabilityC} to collect all the stability related constraints.
% 
\texttt{StabilityC} states that every molecule is moving in a cycle.
\begin{align*}
\texttt{StabilityC} = \texttt{Paths}(r) \land \texttt{Loop}(r) 
\end{align*}
% Boolean variable
% F3 and F2 are used to model these constraints.

@@

\subsubsection{$k$-connectivity constraints}
To check whether $k$-connected is a necessary condition, we remove (drop) $k-1$ edges from the graph and if it
disconnects the graph, and we get an assignment. We have a graph that is not  $k$-connected.
%
Constraints \texttt{Fail}(d,k), \texttt{FReach}(d,r') and \texttt{Connected}(r')
encode the relevant constraints for reachability
in the modified VTS. 
%
The following constraints encode that only
existing edges can be dropped and exactly $k-1$ edges are dropped.
\begin{align*}
\texttt{Fail}(d,k) = & 
\bigwedge\limits_{i,j,q} d_{i,j,q} \limplies e_{i,j,q}  \land 
\sum_{i,j,q} d_{i,j,q} = k-1\\
\end{align*}
We need to define reachability in the modified VTS, therefore we use
a new variable $r'_{i,j}$ to encode reachability from $i$ to $j$.
In the following constraint, we encode $r'_{i,j}$ is true if there is an
edge $(i,\_,j)$ and it is not dropped, or there is a node
$i^{\prime}$ such that, there is an edge $(i,\_,i^{\prime})$ which is
not dropped and $r'_{i',j}$ is true.
\begin{align*}
\texttt{FReach}(d,r') = &  \bigwedge\limits_{i,j}  [\bigvee_{q} (e_{i,j,q} \land  \neg d_{i,j,q}) \lor  (\bigvee_{i' \neq i}  r^{\prime}_{i',j} \land  \bigvee_{q} (e_{i,i',q} \land \neg d_{i,i',q}) ] \limplies r^{\prime}_{i,j}  
\end{align*}
Since we search for disconnected modified VTS,
the following constraint encodes that there are 
unreachable pair of nodes in the underlying undirected graph.
% nodes such that
% there is no path between them .
\begin{align*}
   \texttt{Connected}(r')  = & \bigvee\limits_{i,j} \neg (r^{\prime}_{i,j} \lor r^{\prime}_{j,i})
\end{align*}
We define $\texttt{Connected}(k)$ that says graph is not k-connected.
\begin{align*}
\texttt{Connected}(k) = \texttt{Fail}(d,k) \land \texttt{FReach}(d,r') \land \texttt{Connected}(r') 
% \\
% \texttt{Disconnected}(k) = & \exists d.\;
%         (\texttt{Drop}(d,k) \land \exists r'.\;\texttt{ReachAbove}(d,r')
%                               \land \lnot \texttt{Connected}(r'))
\end{align*}
% We can go down or up using -C \_ option.  \newline

### 2019-04-02 papers/journal/intro.tex
SMT encoding with Constraints Tag

\subsection{The model}
\todo{Ankit: The Rothman model has no intuition; Add couple of lines relating it to VTS; Also, relate it to previous subsection. The connect is lacking.} The VTS model for this paper is inspired by~\cite{shukla2017discovering}. 
%
On the timescales of minutes, we have tried to capture important aspect of Rothman-Schekman-Sudhof (RSS) model~\cite{rothman2002machinery} of vesicle traffic system.
%
The following are the basic component and assumptions of our model. 
\begin{enumerate}
\item A cell is a set of compartments exchanging molecules via vesicles of uniform size.
\item On the timescale of minutes, compartments are neither created nor destroyed.
\item Each compartment is in steady state, gain and loss of the molecules is balance\todo{Ankit: Need intuition and also reasoning why this is okay to add this as an assumption}.
\item We assume that molecules are neither created nor destroyed.
\item Molecules move via vesicles of uniform size.
\item Identical vesicles have identical target compartments.
\item Fusion of vesicles to compartments is driven by specific SNARE pairing.
\item The identities of compartments and vesicles is encoded in their molecules, and identical vesicles fuse with identical target compartments.
\item The activity of a SNARE can be regulated by other molecules present on the same compartment
or vesicle.
\item An active SNARE pair is necessary and sufficient for fusion. 
\end{enumerate}

>>>>

\subsection{The model}
\todo{Ankit: The Rothman model has no intuition; Add couple of lines relating it to VTS; Also, relate it to previous subsection. The connect is lacking.} The VTS model for this paper is inspired by~\cite{shukla2017discovering}. 
%
On the timescales of minutes, we have tried to capture important aspect of Rothman-Schekman-Sudhof (RSS) model~\cite{rothman2002machinery} of vesicle traffic system.
%
The following are the basic component and assumptions of our model. 
\begin{enumerate}
\item A cell is a set of compartments exchanging vesicles.
\item Compartments are neither created nor destroyed.
\item Each compartment is in steady state, gain and loss of the molecules is balance\todo{Ankit: Need intuition and also reasoning why this is okay to add this as an assumption}.
\item Molecules are neither created nor destroyed.
\item Molecules move via vesicles of uniform size.
\item Identical vesicles have identical target compartments.
\item Fusion of vesicles to compartments is driven by specific SNARE pairing.
%\item The identities of compartments and vesicles is encoded in their molecules, and identical vesicles fuse with identical target compartments.
\item The activity of a SNARE can be regulated by other molecules present on the same compartment or vesicle.
\item An active SNARE pair is necessary and sufficient for fusion. 
\end{enumerate}
